SMT-Based Context-Bounded Model Checking for Embedded Systems: Challenges and Future Trends

### 2016-03-25 paper.tex
minor changes

There are several invariant generation algorithms that discover linear and polynomial relations among integer and real variables, in order to provide loop invariants and also to discover the memory ``shape'' in programming languages with pointers~\cite{pips:2013,Henry:2012}. The current literature also reports successful applications of \textit{k}-induction based verification algorithms for hardware and software systems using invariant generation and strengthening, mostly based on interval analysis.

Novel verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant) were proposed~\cite{Gadelha15,Beyer15,Brain15,Rocha15}. Additionally, \textit{k}-induction based verification was also applied to ensure that (restricted) C programs (1) do not contain violations related to data races~\cite{Donaldson10,Kinductor} considering the  Cell BE processor and (2) do respect time constraints, which are specified during the system design phase~\cite{EenS03}. The \textit{k}-induction algorithm is also a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,Sheera00,GrosseLD09}.

@@

%--------------------------------------
\subsection{Incorporating System Model to Automated Verification Procedures}
%--------------------------------------

@@

Indeed, research in software verification is now taking such considerations into account and some schemes already use knowledge about the system to be verified and the underlying hardware. Recently, a verification tool proposed for digital systems, called digital system verifier (DSVerifier)~\cite{dsv_spin2015}, which is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects. Indeed, DSVerifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, transposed forms, etc.) and other implementation restrictions to explore the design space. Ultimately, if the system requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.

@@

State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification, and high-level processor models during simulation~\cite{Behrend15,Lettnin09} as shown in Figure~\ref{verification-methodologies}. 
%
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.35]{figure2.eps}
	\caption{Verification methodologies for embedded systems.}
	\label{verification-methodologies}
\end{figure}

@@

\textbf{(RP1)} Cordeiro, Fischer, and Marques-Silva proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}, CUDA programs~\cite{Pereira15}, and Qt-based consumer electronics applications~\cite{Sousa15}. This approach was also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC)~\cite{CordeiroF11,Cordeiro12}. Other SMT-based BMC approaches have also been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing ones are still limited to specific classes of programs, especially for those that contain intensive floating-point arithmetic and dynamic memory allocation~\cite{Beyer14,BeyerSVCOMP15}. One possible research direction is to bridge the gap between BMC tools and SMT solvers to develop more efficient decision procedures to handle specific class of programs.

\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro, Fischer, and Marques-Silva was further developed to verify correct lock acquisition ordering and the absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}. Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prune the state-space exploration~\cite{Inverso14,civl15}. However, the class of concurrent programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited. One possible research direction is to further extend BMC of multi-threaded C programs via Lazy Sequentialization~\cite{Inverso14}, in order to analyze unsatisfiability cores with the goal to remove redundant behaviour.

@@

\textbf{(RP4)} State-of-the-art SMT-based context-BMC approaches were extended to verify overflow, limit cycle, time constraints, stability, and minimum phase, in digital systems; in particular, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter}, and to specify system-level properties of those systems using linear-time temporal logic~\cite{JMorse15}. In particular, a specific UAV application was tackled, with the goal to verify its attitude controllers~\cite{Bessa16}. In general, however, there is still a lack of studies to verify system-level properties related to embedded systems; emphasis should be given to micro-grids~\cite{xu15} and cyber-physical systems~\cite{leeCPS2}, which require high-dependability requirements for computation, control, and communication. Additionally, the application of automated fault detection, localization, and correction techniques to digital systems represents an important research direction to make BMC tools useful for engineers~\cite{Alves15}.

>>>>

There are several invariant-generation algorithms that discover linear and polynomial relations among integer and real variables, in order to provide loop invariants and also to discover the memory ``shape'' in programming languages with pointers~\cite{pips:2013,Henry:2012}. The current literature also reports successful applications of \textit{k}-induction based verification algorithms for hardware and software systems using invariant generation and strengthening, mostly based on interval analysis.

Novel verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant) were proposed~\cite{Gadelha15,Beyer15,Brain15,Rocha15}. Additionally, \textit{k}-induction based verification was also applied to ensure that (restricted) C programs (1) do not contain violations related to data races~\cite{Donaldson10,Kinductor} considering the  Cell BE processor and (2) do respect time constraints, which are specified during the system design phase~\cite{EenS03}. Apart from that, the \textit{k}-induction algorithm is also a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,Sheera00,GrosseLD09}.

@@

%--------------------------------------
\subsection{Incorporating System Models to Automated Verification Procedures}
%--------------------------------------

@@

Indeed, research in software verification is now taking such considerations into account and some schemes already use knowledge about the system to be verified and the underlying hardware. Recently, a verification tool proposed for digital systems, called digital system verifier (DSVerifier)~\cite{dsv_spin2015}, which is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects. Indeed, DSVerifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, delta forms, and transposed forms) and other implementation restrictions to explore the design space. Ultimately, if the system requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.

@@

State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification, and high-level processor models during simulation~\cite{Behrend15,Lettnin09}, as shown in Figure~\ref{verification-methodologies}. 
%
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.35]{figure2.eps}
	\caption{Verification methodologies for embedded systems.}
	\label{verification-methodologies}
\end{figure}

@@

\textbf{(RP1)} Cordeiro, Fischer, and Marques-Silva proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}, CUDA programs~\cite{Pereira15}, and Qt-based consumer electronics applications~\cite{Sousa15}. This approach was also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC)~\cite{CordeiroF11,Cordeiro12}. Other SMT-based BMC approaches have also been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing ones are still limited to specific classes of programs, especially for those that contain intensive floating-point arithmetic and dynamic memory allocation~\cite{Beyer14,BeyerSVCOMP15}. One possible research direction is to bridge the gap between BMC tools and SMT solvers to propose theories and develop more efficient decision procedures, in order to handle specific classes of programs.

\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro, Fischer, and Marques-Silva was further developed to verify correct lock acquisition ordering and the absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}. Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prune the state-space exploration~\cite{Inverso14,civl15}. However, the class of concurrent programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited. One possible research direction is to further extend BMC of multi-threaded C programs via Lazy Sequentialization~\cite{Inverso14}, in order to analyze unsatisfiability cores with the goal to remove redundant behaviour or to analyze interpolants to prove non-interference of context switches.

@@

\textbf{(RP4)} State-of-the-art SMT-based context-BMC approaches were extended to verify overflow, limit cycle, time constraints, stability, and minimum phase, in digital systems; in particular, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter}, and to specify system-level properties of those systems using linear-time temporal logic~\cite{JMorse15}. In particular, a specific UAV application was tackled, with the goal to verify its attitude controllers~\cite{Bessa16}. In general, however, there is still a lack of studies to verify system-level properties related to embedded systems; emphasis should be given to micro-grids~\cite{xu15} and cyber-physical systems~\cite{leeCPS2}, which require high-dependability requirements for computation, control, and communication. Additionally, the application of automated fault detection, localization, and correction techniques to digital systems represents an important research direction to make BMC tools useful for embedded systems engineers~\cite{Alves15}.
